The Error Estimate of Force Calculation in the Inhomogeneous Molecular Systems

### 2012-03-04 longRange.hetero/publish/main.tex
added staggered mesh section

As mentioned above, the SPME error mainly oringinated from approximate
the term $e^{2\pi i\v m\cdot\v r}$ by the B-spline
interpolation. Denote the error introduced by this approximation by
$A(\v m, \v r)\,e^{2\pi i\v m\cdot \v r}$, we are fortunate to have
the analytical expression of $A(\v m, \v r)$:
\begin{align}\label{eqn:tmp18}
  A(\v m, \v r)
  =
  \sum_{\alpha=1}^3\sum_{l\neq 0}
  \frac
  {\big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}
  {\sum\limits_{l=-\infty}^{+\infty}
    \big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}
  \bigg(
  e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot \v r} - 1
  \bigg).
\end{align}
The error force of the ik-differentiation is therfore
\begin{align}\nonumber
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)
  =\;&
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  A(\v m, \v r_i)\,
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,A(\v m, -\v r_j)\,e^{-2\pi i\v m\cdot\v r_j}\\\label{eqn:tmp19}
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  \big[
  A(\v m, \v r_i) +
  A(\v m, -\v r_i)
  \big].
\end{align}
The last ``self-interacting term'' is zero, because of the
symmetricity of $\v g(\v m)$ and $A(\v m, \v r)$.  Unfortunately, the
error force of ik-differentiation~\eqref{eqn:tmp19} cannot be written
in the form of~\eqref{eqn:tmp11}, so the error estimate derived by
Theorem~\ref{thm:tmp1} cannot be used here. However, it is still
possible to develop the error estimate for the ik-differentiation.
Here we only want to present the key idea and the formula of the error
estimate, and leave all details to the Appendix \ref{}.  We notice the
term $e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot\v r}$ with $l\neq 0$
in Eqn.~\eqref{eqn:tmp18} introduces some high-wave-number terms in
the expression of the error estimate.  If $K_\alpha$ is large enough,
then due to the thermal fluctuation of the system, the peaks and
valleys of the high-wave-number terms may canncels in the the
emsemble-averaged error estimate.
% Based on this observation, it is
% still possible to develop the error estimate for the
% ik-differentiation.
% And if the high wave number terms are neglected,
Therefore, by neglecting these terms, it is proved that the mean error
force of the ik-differentiation is
\begin{align}
  \big\langle
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)
  \big\rangle
  & \approx
  - 2 q_i
  \bigg[
  \big(
  \sum_{\alpha}\sum_{l\neq 0}\v G_{\alpha,l}
  \big)
  \ast\rho_q\,
  \bigg] (\v r_i),
\end{align}
% Also neglect the high wave number terms,
and the RMS force error is 
\begin{align}\nonumber
  \big\langle
  \vert\Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)\vert^2
  \big\rangle
  \approx&\, 
  2q_i^2
  \bigg[\,
  \big(
  \sum_{\alpha} \sum_{l\neq 0}
  \vert \v G_{\alpha,l}\vert^2
  \big)
  \ast \rho_{q^2}
  \,\bigg] (\v r_i)
  +
  4q_i^2\,
  \bigg[
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  \v G_{\alpha,l}
  \big)^2
  \ast \rho_{q^2}\,
  \bigg] (\v r_i) \\
  &\,+
  \big\langle
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)
  \big\rangle^2.
\end{align}
Where
\begin{align}
  \v G_{\alpha,l}(\v r) =
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  \frac
  {\big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}
  {\sum\limits_{l=-\infty}^{+\infty}
    \big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}
  e^{2\pi i\v m\cdot\v r}.
\end{align}
$G_{\alpha,l}(\v r)$ does not depends on the coordinates of particles,
so it can be calculated once and stored in a table for further uses.

@@

The error force of the analytical differetiation is 
\begin{align}\nonumber
  \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)
  =\;&
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \nabla_{\v r_i}
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,
  A(\v m, -\v r_j)
  e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \nabla_{\v r_i}
  [A(\v m, \v r_i)\,
  e^{2\pi i\v m\cdot \v r_i} ]
  \sum_{j\neq i}q_j\,
  e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  & +\,
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\\nonumber
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  A(-\v m, \v r_i)\,e^{2\pi i\v m\cdot \v r_i} \nabla_{\v r_i}e^{2\pi i\v m\cdot\v r_i} \\
  & +\,
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  e^{-2\pi i\v m\cdot \v r_i}
  \nabla_{\v r_i} [\,A(\v m, \v r_i) \,e^{2\pi i\v m\cdot\v r_i}\,]  \\\nonumber
  =&\,
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)\\\nonumber
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \v B(\v m, \v r_i)\,e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,
  e^{-2\pi i\v m\cdot\v r_j} \\\label{eqn:ana.error.force}
  & +\,
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,\v B(\v m, \v r_i) 
\end{align}
With $\v B(\v m, \v r)$ defined by
\begin{align}\nonumber
  \v B(\v m, \v r)
  &=
  \nabla_{\v r_i}A(\v m, \v r)\\
  &=
  \sum_{\alpha=1}^3\sum_{l\neq 0}
  \frac
  {\big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}
  {\sum\limits_{l=-\infty}^{+\infty}
    \big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}
  2\pi i l K_\alpha\v a_\alpha^\ast\,e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot \v r} 
\end{align}
The last term in Eqn.~\eqref{eqn:ana.error.force} is due to the
self-interaction~\cite{cerutti2009staggered, ballenegger2011removal,
  neelov2010interlaced}, which does not depends on the coordinates of
particles. Therefore, it can be calculated explicitly, and substracted
from the analytical force during the simulation. The computational cost of
the self-interacting term costs is very low comparing with the SPME
force calculation.

Neglecting all high frequency terms, we have the mean error force of
analytical differetiation:
\begin{align}\label{eqn:tmp31}
  \langle \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)\rangle
  \approx
  \langle \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)\rangle.
\end{align}
The RMS force error is 
\begin{align}\nonumber
  \big\langle
  \vert \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)\vert^2
  \big\rangle
  \approx&\,
  q_i^2
  \bigg[\,
  \big(
  \sum_{\alpha} \sum_{l\neq 0}
  \vert \v G_{\alpha,l}\vert^2
  \big)
  \ast \rho_{q^2}
  \,\bigg] (\v r_i) \\\nonumber
  & +\,
  q_i^2
  \bigg[
  \big(
  \sum_{\alpha}\sum_{l\neq 0}
  \vert
  \v G_{\alpha,l} + 2\pi i\,l K_\alpha F_{\alpha,l} \v a_\alpha^\ast\vert^2
  \big)
  \ast \rho_{q^2}
  \bigg]
  (\v r_i) \\ \nonumber
  & +\,
  4q_i^2\,
  \bigg[
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  \v G_{\alpha,l}
  \big)^2
  \ast \rho_{q^2}\,
  \bigg] (\v r_i) \\\nonumber
  & +\,
  \big\langle \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)\big\rangle ^2\\
  &+\,
  q_i^4 \sum_{\alpha}\sum_{l\neq 0}
  4\pi^2l^2K_\alpha^2\vert\v a_{\alpha}^\ast\vert^2[C_{\alpha,l}^{\textrm{self}}]^2 .
\end{align}
% The last one is a constant that stems from the self-interaction, which
% do not depends on the
Where
\begin{align}
  F_{\alpha,l} (\v r)
  &=
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  (-2)f(\v m)\,
  \frac
  {\big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}
  {\sum\limits_{l=-\infty}^{+\infty}
    \big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}\,
  e^{2\pi i\v m\cdot \v r},\\
  C_{\alpha,l}^{\textrm{self}}
  &= 
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  (-2)f(\v m)\,
  \frac
  {\big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}
  {\sum\limits_{l=-\infty}^{+\infty}
    \big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}.
\end{align}
$F_{\alpha,l} (\v r)$ and $C_{\alpha,l}^{\textrm{self}}$ 
do not depends on the coordinates of particles,
so they can be tabulated in advance for the simulation.

@@

Fig.~\ref{fig:meanf1} presents the mean error force of the direct part
$\langle\Delta \v F_{\textrm{dir}}\rangle$, reciprocal part of Ewald
summation $\langle\Delta \v F_{\textrm{rec}}\rangle$ and the
reciprocal part of the analytical differentiation $\langle\Delta \v
F^{\textrm{ana}}_{\textrm{rec}}\rangle$. The total mean error force of
analytical differentiation, which is simply $\langle\Delta \v
F^{\textrm{ana}}\rangle = \langle\Delta \v F_{\textrm{dir}}\rangle +
\langle\Delta \v F^{\textrm{ana}}_{\textrm{rec}}\rangle$, is also
given.  The real and estimated mean error forces are denoted by dots
and straight lines, respectively. All estimates are consistent very
well with real errors.  Since the system is locally neutral, the first
order charge distribution $\rho_q$ is naturally zero, therefore, by
Eqn. \eqref{eqn:tmp10}, \eqref{eqn:tmp18} and \eqref{eqn:tmp31}, all
the mean error forces should varnish. Numerical resutls in
Fig. \ref{fig:meanf1} verifies with the above discussion.

>>>>

As mentioned above, the SPME error mainly oringinated from approximate
the term $e^{2\pi i\v m\cdot\v r}$ by the B-spline
interpolation. Denote the error introduced by this approximation by
$A(\v m, \v r)\,e^{2\pi i\v m\cdot \v r}$, we are fortunate to have
the analytical expression of $A(\v m, \v r)$:
\begin{align}\label{eqn:error-A}
  A(\v m, \v r)
  =
  \sum_{\alpha=1}^3\sum_{l\neq 0}
  Z_{\alpha,l}(\v m)
  (
  e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot \v r} - 1
  ),
with
\begin{align}
  Z_{\alpha,l}(\v m) = \frac
  {\big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}
  {\sum\limits_{l=-\infty}^{+\infty}
    \big(\dfrac{2\pi m_\alpha}{K_\alpha} + 2\pi l\big)^{-n}}.
The error force of the ik-differentiation is therfore
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)
  =\;&
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  A(\v m, \v r_i)\,
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,A(\v m, -\v r_j)\,e^{-2\pi i\v m\cdot\v r_j}\\\label{eqn:tmp19}
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  \big[
  A(\v m, \v r_i) +
  A(\v m, -\v r_i)
  \big].
The last ``self-interacting term'' is zero, because of the
symmetricity of $\v g(\v m)$ and $A(\v m, \v r)$.
Therefore,
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)
  =\;&
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  A(\v m, \v r_i)\,
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,e^{-2\pi i\v m\cdot\v r_j} \\\label{eqn:ik.error.force}
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,A(\v m, -\v r_j)\,e^{-2\pi i\v m\cdot\v r_j}
Unfortunately, the
error force of ik-differentiation~\eqref{eqn:ik.error.force} cannot be written
in the form of~\eqref{eqn:tmp11}, so the error estimate derived by
Theorem~\ref{thm:tmp1} cannot be used here. However, it is still
possible to develop the error estimate for the ik-differentiation.
Here we only want to present the key idea and the formula of the error
estimate, and leave all details to the Appendix \ref{}.  We notice the
term $e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot\v r}$ with $l\neq 0$
in Eqn.~\eqref{eqn:error-A} introduces some high-wave-number terms in
the expression of the error estimate.  If $K_\alpha$ is large enough,
then due to the thermal fluctuation of the system, the peaks and
valleys of the high-wave-number terms may canncels in the the
emsemble-averaged error estimate.
% Based on this observation, it is
% still possible to develop the error estimate for the
% ik-differentiation.
% And if the high wave number terms are neglected,
Therefore, by neglecting these terms, it is proved that the mean error
force of the ik-differentiation is
\begin{align}\label{eqn:ik.meanf}
  \big\langle
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)
  \big\rangle
  & \approx
  - 2 q_i
  \bigg[
  \big(
  \sum_{\alpha}\sum_{l\neq 0}\v G_{\alpha,l}
  \big)
  \ast\rho_q\,
  \bigg] (\v r_i),
% Also neglect the high wave number terms,
and the RMS force error is 
  \big\langle
  \vert\Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)\vert^2
  \big\rangle
  \approx&\, 
  2q_i^2
  \bigg[\,
  \big(
  \sum_{\alpha} \sum_{l\neq 0}
  \vert \v G_{\alpha,l}\vert^2
  \big)
  \ast \rho_{q^2}
  \,\bigg] (\v r_i)\\\nonumber
  &\,+
  4q_i^2\,
  \bigg[
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  \v G_{\alpha,l}
  \big)^2
  \ast \rho_{q^2}\,
  \bigg] (\v r_i) \\\label{eqn:ik.error}
  &\,+
  \big\langle
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)
  \big\rangle^2.
Where
\begin{align}
  \v G_{\alpha,l}(\v r) =
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m)\,
  Z_{\alpha,l}(\v m)\,
  e^{2\pi i\v m\cdot\v r}.
$G_{\alpha,l}(\v r)$ does not depends on the coordinates of particles,
so it can be calculated once and stored in a table for further uses.

@@

The error force of the analytical differetiation is 
\begin{align}\nonumber
  \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)
  =\;&
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \nabla_{\v r_i}
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,
  A(\v m, -\v r_j)
  e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \nabla_{\v r_i}
  [A(\v m, \v r_i)\,
  e^{2\pi i\v m\cdot \v r_i} ]
  \sum_{j\neq i}q_j\,
  e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  & +\,
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\\nonumber
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  A(-\v m, \v r_i)\,e^{2\pi i\v m\cdot \v r_i} \nabla_{\v r_i}e^{2\pi i\v m\cdot\v r_i} \\\nonumber
  & +\,
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  e^{-2\pi i\v m\cdot \v r_i}
  \nabla_{\v r_i} [\,A(\v m, \v r_i) \,e^{2\pi i\v m\cdot\v r_i}\,]  \\\nonumber
  =&\,
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)\\\nonumber
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \v B(\v m, \v r_i)\,e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,
  e^{-2\pi i\v m\cdot\v r_j} \\\label{eqn:ana.error.force}
  & +\,
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,\v B(\v m, \v r_i) 
\end{align}
With $\v B(\v m, \v r)$ defined by
\begin{align}\nonumber
  \v B(\v m, \v r)
  &=
  \nabla_{\v r}\,A(\v m, \v r)\\
  &=
  \sum_{\alpha=1}^3\sum_{l\neq 0}
  Z_{\alpha,l}(\v m)\,
  2\pi i l K_\alpha\v a_\alpha^\ast\,e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot \v r} 
\end{align}
The last term in Eqn.~\eqref{eqn:ana.error.force} is due to the
self-interaction~\cite{cerutti2009staggered, ballenegger2011removal,
  neelov2010interlaced}, which does not depends on the coordinates of
particles. Therefore, it can be calculated explicitly, and substracted
from the analytical force during the simulation. The computational cost of
the self-interacting term costs is very low comparing with the SPME
force calculation.

Neglecting all high frequency terms, we have the mean error force of
analytical differetiation:
\begin{align}\label{eqn:ana.meanf}
  \langle \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)\rangle
  \approx
  \langle \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)\rangle.
\end{align}
The RMS force error is 
\begin{align}\nonumber
  \big\langle
  \vert \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)\vert^2
  \big\rangle
  \approx&\,
  q_i^2
  \bigg[\,
  \big(
  \sum_{\alpha} \sum_{l\neq 0}
  \vert \v G_{\alpha,l}\vert^2
  \big)
  \ast \rho_{q^2}
  \,\bigg] (\v r_i) \\\nonumber
  & +\,
  q_i^2
  \bigg[
  \big(
  \sum_{\alpha}\sum_{l\neq 0}
  \vert
  \v G_{\alpha,l} + 2\pi i\,l K_\alpha F_{\alpha,l} \v a_\alpha^\ast\vert^2
  \big)
  \ast \rho_{q^2}
  \bigg]
  (\v r_i) \\ \nonumber
  & +\,
  4q_i^2\,
  \bigg[
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  \v G_{\alpha,l}
  \big)^2
  \ast \rho_{q^2}\,
  \bigg] (\v r_i) \\\nonumber
  & +\,
  \big\langle \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)\big\rangle ^2\\\label{eqn:ana.error}
  &+\,
  q_i^4 \sum_{\alpha}\sum_{l\neq 0}
  4\pi^2l^2K_\alpha^2\vert\v a_{\alpha}^\ast\vert^2[C_{\alpha,l}^{\textrm{self}}]^2 .
\end{align}
% The last one is a constant that stems from the self-interaction, which
% do not depends on the
Where
\begin{align}
  F_{\alpha,l} (\v r)
  &=
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  (-2)f(\v m)\,
  Z_{\alpha,l}(\v m)\,
  e^{2\pi i\v m\cdot \v r},\\
  C_{\alpha,l}^{\textrm{self}}
  &= 
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  (-2)f(\v m)\,
  Z_{\alpha,l}(\v m).
\end{align}
$F_{\alpha,l} (\v r)$ and $C_{\alpha,l}^{\textrm{self}}$ 
do not depends on the coordinates of particles,
so they can be tabulated in advance for the simulation.

\subsection{The staggered Mesh Ewald method}

The staggered Method Ewald method~\cite{cerutti2009staggered,
  neelov2010interlaced} calculate the reciprocal force on two meshes,
one of which locates at the middle of the other.  Then the two forces
are averaged to cancel the error. Here we study the staggered Mesh
Ewald method from the error point of view.
The shift of the mesh can be effective treated as shifting all
particles on the opposite direction. Since the system is periodic, the
shift to which direction is not important. For convenience, we shift
all particles to the positive direction for half the grid size, namely
by a vector $\v s = \sum_\alpha\frac1{2K_\alpha}\v a_\alpha$. From
Eqn.~\eqref{eqn:error-A}, we have
\begin{align}
  \frac12\, [A(\v m, \v r) + A(\v m, \v r+\v s)]
  =&
  \sum_{\alpha=1}^3\sum_{l\neq 0}
  \frac12\,
  Z_{\alpha,l}(\v m)\,
  \bigg[
  (1 + e^{\pi i l})e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot \v r} - 2
  \bigg].
\end{align}
Therefore, when $l$ is odd, $1 + e^{\pi i l}$ vanishes. So the $\frac12\,[A(\v
m, \v r) + A(\v m, \v r+\v s)]$ can be approximated by the leading
order:
\begin{align}
  \frac12\, [A(\v m, \v r) + A(\v m, \v r+\v s)]
  \approx&
  -\sum_{\alpha=1}^3\sum_{l\neq 0}
  Z_{\alpha,l}(\v m).
\end{align}
Then from Eqn.~\eqref{eqn:ik.error.force}, we have the error force of
the staggered ik-differentiation:
\begin{align}
  \Delta\v F^{\textrm{ik,st}}_{\textrm{rec}}(\v r_i)
  \approx -2q_i
  \sum_jq_j
  \sum_\alpha\sum_{l\neq 0}
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  g(\v m) Z_{\alpha,l}(\v m)\,e^{2\pi i\v m\cdot(\v r_i-\v r_j)}.
\end{align}
This is a pairwise interaction, therefore, from
Theorem~\ref{thm:tmp1}, we have the error estimates:
\begin{align}\label{eqn:ik.st.meanf}
  \big\langle
  \Delta\v F_{\textrm{rec}}^{\textrm{ik,st}}(\v r_i)
  \big\rangle
  & \approx
  - 2 q_i
  \bigg[
  \big(
  \sum_{\alpha}\sum_{l\neq 0}\v G_{\alpha,l}
  \big)
  \ast\rho_q\,
  \bigg] (\v r_i),
\end{align}
% Also neglect the high wave number terms,
and the RMS force error is 
\begin{align}\label{eqn:ik.st.error}
  \big\langle
  \vert\Delta\v F_{\textrm{rec}}^{\textrm{ik,st}}(\v r_i)\vert^2
  \big\rangle
  \approx&\,
  4q_i^2\,
  \bigg[
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  \v G_{\alpha,l}
  \big)^2
  \ast \rho_{q^2}\,
  \bigg] (\v r_i) +
  \big\langle
  \Delta\v F_{\textrm{rec}}^{\textrm{ik,st}}(\v r_i)
  \big\rangle^2.
\end{align}
The superscript ``st'' denotes the staggered mesh method.
The mean error force of the staggered ik-differentiation is the
same as the staggered ik-differentiation. The RMS error is improved, because
$\big\langle\vert\Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)\vert^2\big\rangle -
\big\langle\vert\Delta\v F_{\textrm{rec}}^{\textrm{ik,st}}(\v r_i)\vert^2\big\rangle=
2q_i^2
\big[\,
\big(
\sum_{\alpha} \sum_{l\neq 0}
\vert \v G_{\alpha,l}\vert^2
\big)
\ast \rho_{q^2}
\,\big] (\v r_i) \geq 0$.

For the analytical differetiation, similarly we have: $\frac12\,[\v
B(\v m,\v r) + \v B(\v m, \v r+\v s)] \approx 0$. Therefore, from
Eqn.~\eqref{eqn:ana.error.force}, the error force of the staggered
analytical differetiation is approximately the same as the staggered
ik-differentiation, so are all the error estimates:
\begin{align}
  \Delta\v F^{\textrm{ana,st}}_{\textrm{rec}}(\v r_i)
  &\approx
  \Delta\v F^{\textrm{ik,st}}_{\textrm{rec}}(\v r_i)\\
  \big\langle
  \Delta\v F_{\textrm{rec}}^{\textrm{ana,st}}(\v r_i)
  \big\rangle
  &\approx
  \big\langle
  \Delta\v F_{\textrm{rec}}^{\textrm{ik,st}}(\v r_i)
  \big\rangle\\
  \big\langle
  \vert\Delta\v F_{\textrm{rec}}^{\textrm{ana,st}}(\v r_i)\vert^2
  \big\rangle
  &\approx
  \big\langle
  \vert\Delta\v F_{\textrm{rec}}^{\textrm{ik,st}}(\v r_i)\vert^2
  \big\rangle
\end{align}
The error from the self-interaction term is automatically fixed by the
staggered mesh. The RMS force error is improved because it is not
difficult to see 
$\big\langle\vert\Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)\vert^2\big\rangle -
\big\langle\vert\Delta\v F_{\textrm{rec}}^{\textrm{ana,st}}(\v r_i)\vert^2\big\rangle
\geq 0$.

@@

Fig.~\ref{fig:meanf1} presents the mean error force of the direct part
$\langle\Delta \v F_{\textrm{dir}}\rangle$, reciprocal part of Ewald
summation $\langle\Delta \v F_{\textrm{rec}}\rangle$ and the
reciprocal part of the analytical differentiation $\langle\Delta \v
F^{\textrm{ana}}_{\textrm{rec}}\rangle$. The total mean error force of
analytical differentiation, which is simply $\langle\Delta \v
F^{\textrm{ana}}\rangle = \langle\Delta \v F_{\textrm{dir}}\rangle +
\langle\Delta \v F^{\textrm{ana}}_{\textrm{rec}}\rangle$, is also
given.  The real and estimated mean error forces are denoted by dots
and straight lines, respectively. All estimates are consistent very
well with real errors.  Since the system is locally neutral, the first
order charge distribution $\rho_q$ is naturally zero, therefore, by
Eqn. \eqref{eqn:tmp10}, \eqref{eqn:tmp18} and \eqref{eqn:ana.meanf}, all
the mean error forces should varnish. Numerical resutls in
Fig. \ref{fig:meanf1} verifies with the above discussion.
